Paper_Notes

### 2017-11-24 Paper_Notes.Rmd
Reading Assignment for 11/20/17 and 11/27/17

- Cleveland, William S, and Susan J Devlin. "Locally Weighted Regression: An Approach to Regression Analysis by Local Fitting." Journal of the American Statistical Association 83, no. 403 (1988): 596-610. - Loess regression. Concept, statistics.

@@

>>>>

- Cleveland and Devlin. "Locally Weighted Regression: An Approach to Regression Analysis by Local Fitting." Journal of the American Statistical Association 83, no. 403 (1988): 596-610.

@@

### Tusher et al. 2001

- Tusher, Tibshirani, and Chu. "Significance analysis of microarrays applied to the ionizing radiation response." PNAS 98, no. 9 (2001): 5116 - 21.

In this article, the authors describe a method, Significance Analysis of Microarrays (SAM), which determines the significance of changes in gene expression between different biological states determined by microarray while accounting for the enormous number of genes measured. Methods based on conventional t tests provide the probability that a difference in gene expression occurred by chance. Although P= 0.01 is significant in the context of experiments designed to evaluate small numbers of genes, a microarray experiment for 10,000 genes would identify 100 genes by chance which is problematic. The authors developed SAM, which identifies genes with statistically significant changes in expression by assimilating a set of gene-specific t tests. Each gene is assigned a score based on its change in gene expression relative to the standard deviation of repeated measurements for that gene. For genes with scores greater than an adjustable threshold, SAM uses permutations of the repeated measurements to estimate the percentage of genes identified by chance, which is the false discovery rate (FDR). During multiple testing, the family-wise error rate is the probability of at least one false positive over the collection of tests. SAM does not have a strong or weak control over the family-wise error rate, and, instead, provides an estimate of the FDR for each value of the tuning parameter define within the algorithm. The authors deem it plausible that the estimated FDR approximates the strongly controlled FDR when any subset of null hypotheses is true. The article demonstrates the efficacy of SAM by assessing the transcriptional response of lymphoblastoid cells to ionizing radiation. The results suggest that SAM is a robust and straightforward method that can be adapted to a broad range of experimental situations.

### Smyth 2004

- Smyth. “Linear Models and Empirical Bayes Methods for Assessing Differential Expression in Microarray Experiments.” Statistical Applications in Genetics and Molecular Biology 3, no. 1 (2004): 1 – 25.

Smyth describes the problem of identifying differentially expressed genes in designed microarray experiments. The massive amount of data assessed in microarray experiments results in a multiple testing problem in which one or more tests are conducted for each of tens of thousands of genes. This is then complicated by the fact that the measured expression levels are often non-normally distributed and have non-identical and dependent distributions between genes. The author describes striking a balance between controlling the familywise error rate or false discovery rate (FDR) and the parallel nature of the inference in microarrays, which allows some compensating possibilities for borrowing information from the ensemble of genes, which can assist in inference about each gene individually. Previously, an article described and derived a method that can reconcile this issue through the application of Bayes or empirical Bayes methods. Using a replicated two-color microarray experiment, Lonnstedt and Speed (2002), took a parametric empirical Bayes approach and derived a simple expression for the posterior odds of differential expression for each gene. The model is reset in the context of general linear models with arbitrary coefficients and contrasts of interest while closed form estimators are derived for the hyperparameters in the model. Furthermore, this approach shrinks the estimated sample variance towards a pooled estimate, resulting in far more stable inference when the number of arrays is small. The method is demonstrated for simulated data, and the results indicate that the proposed estimators have robust behavior that allow for incomplete data that can sometimes arise from spot filtering or spot quality weights.

### Benjamini et al. 1995

- Benjamini and Hochberg. “Controlling the False Discovery Rate: a Practical and Powerful Approach to Multiple Testing.” J.R. Statist. Soc. B 57, no. 1 (1995): 289 – 300.

The authors describe a new point of view on the problem of the multiplicity (selection) effect. In many multiplicity problems, the number of erroneous rejections should be taken into account. Far too often only the question of whether any error was made, which can lead to a serious loss incurred by erroneous rejections. The common approach to the multiplicity problem calls for controlling the familywise error rate (FWER). However, the article points out many issues that arise when using methods that control the FWER including: 1) application of multiple-treatment type tests to non-multiple treatment experiments; 2) applications of these tests tend to substantially reduce the power than the per comparison procedure of the same level; and 3) redundant application of FWER control when it is not needed. Instead, the authors describe a different approach that calls for controlling the expected proportion of falsely rejected hypotheses. This false discovery rate (FDR) is equivalent to the FWER when all hypotheses are true but is smaller otherwise. This leads to potential gain in power when controlling the FDR as opposed to the FWER. In many applications, this is the desirable control against errors originating from multiplicity. A simple sequential Bonferroni-type procedure is shown to control the FDR for independent test statistics. The results demonstrate that controlling the FDR is both simple and powerful; however, it does not take into the structure of specific problems such as pairwise comparisons in analysis of variance.

### Storey et al. 2003

- Storey and Tibshirani. “Statistical Significance for Genomewide Studies.” PNAS 100, no. 16 (2003): 9440 – 45.

Genomewide experiments and the sequencing of multiple genomes have become commonplace, which consequently has raised issues pertaining to the analyses of large data sets. The analyses of these data also involve performing statistical tests on thousands of features in a genome. Prior to performing the analysis, it is expected that many more than one or two of the tested features will be statistically significant. Guarding against any single false positive occurring is often going to be much too strict and will lead to many missed findings. The authors propose an approach to measuring statistical significance in these genomewide studies based on the concept of the false discovery rate (FDR). The goal is to propose and estimate a measure of significance for each feature that meets the practical goals of the genomewide study that is easily interpreted in terms of the simultaneous testing of thousands of features. Instead of a p-value, a q-value is associated with each tested feature. The q-value is similar to the p-value, except it is a measure of significance in terms of the FDR rather than the false positive rate. The results show that the FDR is a sensible measure of the balance between the number of true positives and false positives. Furthermore, it avoids over-inflation of false positive results, while still offering a more sensible criterion than what has been used previously for linkage studies involving genome scans.
